---
title: "Improve your plots with `ggplot2`"
output: 
  html_notebook:
    code_folding: show
    toc: true
    theme: journal
---

# The dataset

```{r, eval=FALSE}
# Install devtools
# install.packages("devtools")

# Install the package tvseries
# devtools::install_github("mireia-bioinfo/tvseries")
```

For this workshop wer are goint to use the dataset `tvseries_must_watch` from the `tvseries` R package. This dataset contins IMDb ratings for 4 TV series that I personally recommend you should watch! This TV series are: `c("Jane the Virgin", "Crazy Ex-Girlfriend", "Brooklyn Nine-Nine", "The Good Place")`. 

**Warning!** Do not load the `tvseries` library as it will take a while (it has some big datasets). Just assign the `tvseries_must_watch` dataset to a new variable as shown below:

```{r}
# df <- tvseries::tvseries_must_watch 
load("data/tvseries_must_watch.rda")
df <- tvseries_must_watch
df <- df[!is.na(df$average_rating),] # Remove rows with NA average rating
head(df)
```

Here you have some description of dataset columns:

Name             | Description
-----------------|------------------------------------
`tvseries_title` | Name of the TV series.
`episode_title`  | Name of the episode.
`season_number`  | Season number.
`episode_number` | Episode number.
`start_year`     | Year the show started.
`end_year`       | Year the show ended.
`runtime_min`    | Length of the episode in minutes.
`genres`         | Genres of the TV series.
`average_rating` | Average rating of the episode.
`num_votes`      | Number of votes for that episode.

# Using `ggplot2`

```{r install-packages, eval=FALSE}
install.packages("ggplot2", "scales", "palettetown", "ggsci")
```


```{r}
library(ggplot2)
```

## Basic `ggplot2`: Geometries

### **Example 1**: Plot the number of votes for each TV series using boxplots

```{r example-1}
ggplot(df,
       aes(tvseries_title, num_votes)) +
  geom_boxplot()
```

### **Example 2**: Plot the number of votes per season for each TV series using boxplots

```{r example-2}
ggplot(df,
       aes(tvseries_title, num_votes, 
           fill=season_number)) +
  geom_boxplot()
```

```{r example-2-fixed}
ggplot(df,
       aes(tvseries_title, num_votes, 
           fill=factor(season_number))) +
  geom_boxplot()
```


### **Exercise 1**: Plot the average ratings for each TV series using boxplots

```{r exercise-1}
ggplot(df,
       aes(tvseries_title, average_rating)) +
  geom_boxplot()
```

### **Exercise 2**: Plot the average ratings per season for each TV series using boxplots

```{r exercise-2}
ggplot(df,
       aes(tvseries_title, average_rating, 
            fill=factor(season_number))) +
  geom_boxplot()
```

### **Exercise 3**: Plot the distribution of average ratings for each TV series using violin plots and boxplots

```{r exercise-3}
ggplot(df,
       aes(tvseries_title, average_rating)) +
  geom_violin() +
  geom_boxplot(width=.2)
```

### **Example 3**: Plot the number of episodes for each TV series using barplots

```{r example-3-a}
ggplot(df,
       aes(tvseries_title,
           fill=factor(season_number))) +
  geom_bar(stat="count", color="black")
```

```{r example-3-b}
ggplot(df,
       aes(tvseries_title,
           fill=factor(season_number))) +
  geom_bar(stat="count", color="black", 
           position="fill")
```

```{r example-3-c}
ggplot(df,
       aes(tvseries_title,
           fill=factor(season_number))) +
  geom_bar(stat="count", color="black", 
           position="dodge")
```

### **Exercise 4**: Plot the density of average ratings for each TV series.

**Tip**: `geom_density()` is another example of a function that does not need a y aesthetic. It automatically computes and draws the kernel density estimate

```{r exercise-4}
ggplot(df,
       aes(average_rating)) +
  geom_density(aes(fill=tvseries_title),
                   alpha=0.5)
```

## Basic `ggplot2`: Scales

### **Example 4**: This plot looks quite sad and wrong... Let's improve it by tweaking its scales!

```{r example-4-wrong}
ggplot(df,
       aes(tvseries_title, average_rating, 
           color=factor(season_number))) +
  geom_jitter() +
  geom_boxplot()
```

- Labels for the axis and legend are terrible.
- Default `ggplot2` colors.
- Difficult to interpret... Many different boxplot and data points.

```{r example-4-fixed}
ggplot(df,
       aes(tvseries_title, average_rating)) +
  geom_jitter(aes(color=factor(season_number)),
              width=0.2, alpha=0.8) +
  geom_boxplot(fill=NA, width=0.5, outlier.shape=NA, 
               size=1) +
  scale_color_brewer(palette="RdPu", 
                     name="Season", 
                     labels=function(x) paste0("S0", x)) +
  scale_y_continuous(name="Average Rating (IMDb)",
                     breaks=scales::pretty_breaks(),
                     limits=c(0,10)) +
  xlab("TV series title")
```

### **Exercise 5**: Yet another sad plot... Try to improve it as much as you can!

```{r exercise-5-wrong}
ggplot(df,
       aes(runtime_min, fill=tvseries_title)) +
  geom_density()
```

```{r exercise-5-fixed}
ggplot(df,
       aes(runtime_min, fill=tvseries_title)) +
  geom_density(alpha=0.6) +
  ggsci::scale_fill_futurama(name="TV series") +
  scale_x_continuous(name="Episode duration",
                     breaks=scales::pretty_breaks(n=4),
                     labels=function(x) paste(x, 
                                              "min"),
                     limits=c(0, NA)) +
  ylab("Density")
```

## Advanced `ggplot2`
### **Example 5**: Plot the mean average ratings per season for each TV series (`stat_summary()`).

```{r example-5}
ggplot(df,
       aes(factor(season_number), average_rating,
           color=tvseries_title, 
           fill=tvseries_title)) +
  geom_jitter(width=0.2, alpha=0.5) +
  stat_summary(aes(group=tvseries_title), 
               fun.y=mean, geom="line",
               size=1) +
  stat_summary(fun.y=mean, geom="point", 
               size=4, pch=21, color="black") +
  palettetown::scale_color_poke(pokemon="charizard", 
                                name="TV series") +
  palettetown::scale_fill_poke(pokemon="charizard",
                               name="TV series") +
  xlab("Season number") + ylab("Average rating")
```

**Tips**:

- `ggplot2` will try to collapse your legends: if you modify the colors or names, make sure they match if you want a single legend!
- When using `geom_line()` you will generally need to add a grouping for the variables. If not, you will get an error saying `geom_path: Each group consists of only one observation. Do you need to adjust the group aesthetic?`.

### **Example 6**: Zoom into an specific region of the plot (`coord_cartesian()`).

```{r example-6}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=tvseries_title)) +
  scale_color_brewer(palette="Set1", 
                     name="TV series") +
  xlab("# of votes") + ylab("Average rating")
```

```{r example-6-zoom}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=tvseries_title)) +
  scale_color_brewer(palette="Set1", 
                     name="TV series") +
  xlab("# of votes") + ylab("Average rating") +
  coord_cartesian(xlim=c(0,500), ylim=c(7,9))
```


### **Example 7**: Add classification according to the TV series title (`facet_wrap()`)

```{r example-7}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=factor(season_number))) +
  scale_color_brewer(palette="Set1", 
                     name="Season") +
  xlab("# of votes") + ylab("Average rating") +
  ylim(0,10)
```

```{r example-7-facet}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=factor(season_number))) +
  scale_color_brewer(palette="Set1", 
                     name="Season") +
  xlab("# of votes") + ylab("Average rating") +
  ylim(0,10) +
  facet_wrap(~tvseries_title, scales="free_x")
```

### **Example 8**: Replace the default `ggplot2` theme for another one!

```{r example-8,}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=tvseries_title)) +
  scale_color_brewer(palette="Set1", 
                     name="TV series") +
  xlab("# of votes") + ylab("Average rating")
```

```{r example-8-corrected}
ggplot(df,
       aes(num_votes, average_rating)) +
  geom_point(aes(color=tvseries_title)) +
  scale_color_brewer(palette="Set1", 
                     name="TV series") +
  xlab("# of votes") + ylab("Average rating") +
  theme_dark()
```

### **Example 9**: Modify specific `theme()` elements so the plot looks nicer!

```{r example-9}
ggplot(df,
       aes(tvseries_title, average_rating)) +
  geom_jitter(aes(color=factor(season_number)),
              width=0.2, alpha=0.8) +
  geom_boxplot(fill=NA, width=0.5, outlier.shape=NA, 
               size=1) +
  scale_color_brewer(palette="RdPu", 
                     name="Season", 
                     labels=function(x) paste0("S0", 
                                               x)) +
  scale_y_continuous(name="Average Rating (IMDb)",
                     breaks=scales::pretty_breaks(),
                     limits=c(0,10)) +
  xlab("TV series title")
```

```{r example-9-corrected}
ggplot(df,
       aes(tvseries_title, average_rating)) +
  geom_jitter(aes(color=factor(season_number)),
              width=0.2, alpha=0.8) +
  geom_boxplot(fill=NA, width=0.5, outlier.shape=NA, 
               size=1) +
  scale_color_brewer(palette="RdPu", 
                     name="Season", 
                     labels=function(x) paste0("S0", 
                                               x)) +
  scale_y_continuous(name="Average Rating (IMDb)",
                     breaks=scales::pretty_breaks(),
                     limits=c(0,10)) +
  xlab("TV series title") +
  theme(axis.text.x=element_text(angle=30, hjust=1),
        panel.background=element_rect(fill="white"),
        axis.line=element_line(size=0.5),
        legend.position="top",
        axis.text=element_text(size=12),
        axis.title=element_text(size=14))
```


# Expanding `ggplot2`

## Combined plots with `cowplot`

```{r}
# install.packages("cowplot")
library(cowplot)

```

## Animated plots with `gganimate` 

```{r}
# install.packages("gganimate")
library(gganimate)

```

## Interactive plots with `ggiraph`

```{r}
# install.packages("ggiraph")
library(ggiraph)
```



